---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fec16

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/fec16)](https://CRAN.R-project.org/package=fec16)
<!-- badges: end -->


`fec16` contains candidate, committee and candidate-committtee linkage data for the United States 2015-2016 election cycle. To get an understanding of what leads to a win, we have various **a selection of x contributions??** contribution data along with state data of the election winners. 


## Installation

`fec16` is hosted on GitHub and call be installed by running the following:

``` r
devtools::install_github("ranawg/fec16")
```

``` r
library(fec16)
```


## Data

`candidates`: all candidates registered with the FEC during the 2015-2016 election cycle

`committees`: all committees registered with the FEC during the 2015-2016 election cycle

`results`: the results of the 2016 general presidential election

`committee_contributions`: total contributions, aggregated by candidate, from committees

`linkage`: provides linkage ID's for candidates and committees


## Example

`fec16` can be used to summarise data in order see how many candidates are running for elections (in all offices) for the two major parties:
```{r example, warning = FALSE}
library(fec16)
library(tidyverse)

candidates %>% filter(cand_pty_aff == "REP"|cand_pty_aff =="DEM") %>% group_by(cand_pty_aff) %>% summarise(size = n())
```

A data wrangling example that uses two of the data frames could be:
```{r}
cand_cmte <- full_join(candidates, committees, by = "cand_id") %>% filter(cand_pty_aff == "REP"|cand_pty_aff =="DEM") %>% group_by(cand_pty_aff, committee_type) %>% summarise(n = n()) %>% drop_na(committee_type)
cand_cmte
```

And extending that to create a visualisation to see the results easily.
```{r}
ggplot(cand_cmte, aes(x = committee_type, y = n, fill = cand_pty_aff)) + geom_col(position = "dodge") +
  labs(title = "Bar Chart of Total Committees by Type and Party", x = "Committee Type", y = "Count", fill = "Candidate Party Affiliation")
```


Create a scatter-plot of total contributions per candidate, colored by party (only including Democrat and Republican)
```{r message = FALSE}

joined_data <- candidates%>%
  full_join(committee_contributions, by = "cand_id") %>%
  filter(!is.na(total_contributions)) %>%
  filter(cand_pty_aff=="REP" | cand_pty_aff == "DEM") %>%
  filter(total_contributions <= 1e+08)
  
  p <- ggplot(joined_data, aes(x=cand_id, y=total_contributions, color = cand_pty_aff)) +
  geom_point() +
  geom_jitter() +
  theme(axis.text.x=element_blank())
  
p

```


Visualize the results of the elections and see how many poeple voted: 
```{r}
library(ggplot2)
library(dplyr)
library(scales)

results_by_cand <- results %>% 
  drop_na(general_results, fec_id) %>%
 group_by(fec_id, last_name) %>%
  summarise(sum_votes = sum(general_results)) %>% 
  filter( sum_votes >100000)
  
  
 
ggplot(results_by_cand, mapping = aes(x = last_name, y = sum_votes) ) + 
  geom_col() + 
  xlab("Candidates") + 
  ylab("Number of Votes") +
  scale_y_continuous(labels = comma)
```

