---
title: "fec16"
author: "Irene Ryan, Marium Tapal, Rana Ghawagy"
date: "` r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{fec16}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(fec16)
```

The `fec16` packages is a data set package that provides tamed dataframes of Federal elections of the 2015-2016 cycle and a documnet for each dataframe. 



## Why tame data ? 
We want to create a package that is easy to use for people just beginning to learn R such as introductory statistics and data science classes. If the data is being used for teaching, students and instructors do not need to worry about the unnecessary data wrangling and can immediately work on applying new knowledge and analysis without going back to the basics. Our package is inspired by both `nycflights13` by [Hadley](https://github.com/hadley/nycflights13) and `fivethirtyeight` by [Albert Kim](https://github.com/rudeboybert/fivethirtyeight).


### Basics of tame data 

We followed the [tame data principles](https://escholarship.org/uc/item/0rx1231m#article_main) in order to produce uniform dataframes that are easy to navigate and link. We used the following guidelines to tame our data:

1. Variable names: 
  a) Lower case and with underscores instead of spaces using `clean_names()`.
  b) 20 characters or less. 
  c) The same names were used for the same variables in different dataframes. 
  d) Names have a meaningful name. 
2. Variable types:
  a) Encode varaible with dates with `as.Date`, unless it is only a year then it would be numeric
  b) Encode categorical variables that a have a limited number of values with `as.factor` otherwise it would      be character.
3. Tidy data format or long instead of wide.
4. Missing data has value `NA`.
5. Removed signs such as % and $ from numeric values 

### Other notes

-  We edropped varibles that are not useful for analysis from the original FEC datasets. 
-  We Summerised large data sets with relavent data by the candidate. 

## Who should use this package? 

Anyone interested in US politics and elections and want to think critically about the data and what can be inferred. We made this package with students and their instructors in mind to make working with the data itself as smooth as possible.This package would provide inspiration for question and projects that are data-driven and based on the real world. 

Here are some examples for possible questions or projects to invistigate: 
-
-
-

### Examples: 





